An Open and Continuously Updated Fern Tree of Life (FTOL)

### 2022-06-21 ms/manuscript.Rmd
Extract stats for best-scoring tree

@@

The `r extract_iqtree_mod(plastome_iqtree_log)` model was selected according to BIC for both the plastome (backbone) and Sanger analyses.
Two of the ten runs converged (correlation coefficient of split occurrence frequencies >0.99), but the run with the highest log-likelihood, from which the final tree was selected, did not converge (correlation coefficient 0.977 after 1,000 iterations).
The consensus tree had higher log-likelihood (-1183776.726) than the ML tree (-1183911.595), so we only present the topology and divergence times estimated from the consensus tree.

>>>>

```{r phylogeny-stats-log}
sanger_log_files <- fs::dir_ls(
 here::here("ftol_cache/user/intermediates/iqtree/"),
 recurse = TRUE,
 regexp = "sanger_[0-9]+/sanger_alignment.phy.log"
 )

rename_sanger <- function(x) {
  names(x) <- str_match(names(x), "sanger_[0-9]") %>%
    magrittr::extract(,1)
  x
}

sanger_ml_log_rep <- sanger_log_files %>%
  map(read_lines) %>%
  unlist()

ml_scores <- sanger_ml_log_rep[
  str_detect(sanger_ml_log_rep, "BEST SCORE FOUND")] %>%
  rename_sanger() %>%
  map_dbl(parse_number)

con_scores <- sanger_ml_log_rep[
  str_detect(sanger_ml_log_rep, "Log-likelihood of consensus tree")] %>%
  rename_sanger() %>%
  map_chr(~str_remove_all(., "Log-likelihood of consensus tree: ")) %>%
  map_dbl(parse_number)
combined_scores <- ml_scores + con_scores

assertthat::assert_that(
  !any(is.na(ml_scores))
)
assertthat::assert_that(
  !any(is.na(con_scores))
)
assertthat::assert_that(
  length(ml_scores) == length(con_scores),
  msg = "Length of ML and consensus scores don't match"
)

best_index <- which(combined_scores == max(combined_scores))
best_ml_score <- ml_scores[best_index]
best_con_score <- con_scores[best_index]

assertthat::assert_that(
  best_ml_score < best_con_score,
  msg = "Best ML score not less than best consensus score"
)

# check model selected
sanger_models <- sanger_log_files %>%
  map(read_lines) %>%
  map(extract_iqtree_mod) %>%
  map(unique) %>%
  rename_sanger()

assertthat::assert_that(
  length(unique(sanger_models)) == 1,
  msg = "More than one sanger model selected"
)
assertthat::assert_that(
  extract_iqtree_mod(plastome_iqtree_log) == unique(sanger_models),
  msg = "Model selected for plastome and sanger don't match"
)

# check number of runs converged
no_converge <- sanger_ml_log_rep[
  str_detect(sanger_ml_log_rep, "bootstrap analysis did not converge")] %>%
  rename_sanger()
n_converge <- length(sanger_log_files) - length(no_converge) %>%
  english::english()

assertthat::assert_that(
  names(best_index) %in% names(no_converge),
  msg = "Best scoring tree *did* converge")
```

@@

The `r extract_iqtree_mod(plastome_iqtree_log)` model was selected according to BIC for both the plastome (backbone) and Sanger analyses.
`r str_to_sentence(n_converge)` of the ten runs converged (correlation coefficient of split occurrence frequencies >0.99), but the run with the highest log-likelihood, from which the final tree was selected, did not converge (correlation coefficient 0.976 after `r number(1000)` iterations).
The consensus tree had higher log-likelihood (`r best_con_score`) than the ML tree (`r best_ml_score`), so we only present the topology and divergence times estimated from the consensus tree.
